Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-08-11 writing/embedded_meta_learning_main.tex
per CMoS, figure references not capitalized

Given this, the task of mapping inputs to outputs can be framed as trying to find a transformation of the representational space that takes the (embedded) inputs from the training set to embeddings that will decode to the target outputs. These transformations are performed by a system with the following components (see Fig. \ref{architecture_inference_fig}): $\mathcal{M}: \{(Z, Z), ...\} \rightarrow Z $ -- the meta network, which collapses a dataset of (input embedding, target embedding) pairs to produce a single function embedding. $\mathcal{H}: Z \rightarrow \text{parameters}$ -- the hyper network, which maps a function embedding to parameters. $F: Z \rightarrow Z$ -- the transformation, implemented by a deep network parameterized by $\mathcal{H}$. \par

@@

%% Cut example if necessary
As a proof of concept, we first evaluated the system on the task of learning polynomials of degree $\leq 2$ in 4 variables (i.e. the task was to regress functions of the form $p: \mathbb{R}^4 \rightarrow \mathbb{R}$ where $p \in \mathcal{P}_2 \left(\mathbb{R}\right)$, though the model was given no prior inductive bias toward polynomial forms). For example, if $p(w,x,y,z) = x$, the model might see examples like $(-1,1,1,1; 1)$ and $(0.7, 2.1, 1.3, -4; 2.1)$, and be evaluated on its output for points like $(-1, -1.3, 0.5, 0.3)$. This yields an infinite family of base-level tasks (the vector space of all such polynomials), as well as many families of meta-mappings over tasks (for example, multiplying polynomials by a constant, squaring them, or permuting their input variables). This allows us to not only examine the ability of the system to learn to learn polynomials from data, but also to adapt its learned representations in accordance with these meta-tasks. Details of the architecture and training can be found in appendix \ref{app_detailed_methods}.\par
\vspace{-0.7em}
\paragraph{Basic meta-learning:} First, we show that the system is able to achieve the basic goal of learning a held-out polynomial from a few data points in Fig. \ref{poly_basic_results} (with good sample-efficiency, see supp. Fig. \ref{supp_poly_sweep_results}). \par 
\vspace{-0.7em}
\paragraph{Meta-mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings over polynomials in order to flexibly reconfigure its behavior, as shown in Fig. \ref{poly_meta_map_results}. We train the system to perform a variety of mappings, for example switch the first two inputs of the polynomial, add 3 to the polynomial, or square the polynomial. We then test its ability to generalize to held-out mappings, for example a permutation of the polynomial inputs which it has not encountered before, or an additive shift in the polynomial which it has not been trained on. The system is both able to apply its learned meta-mappings to held-out tasks, and to apply meta-mappings it has not been trained on, simply by seeing examples of the mapping. \par

@@

\paragraph{Basic meta-learning:} First, we show that the system is able to achieve the basic goal of playing a held-out game from examples in Fig. \ref{cards_basic_results}. We compare two different possible hold-out sets: 1) train on half the tasks at random, or 2) specifically hold out all the ``losers'' variations of the ``straight flush'' game. In either of these cases, the meta-learning system achieves well above chance performance (0) at the held out tasks, although it is slightly worse at generalizing to the targeted hold out, despite having more training tasks in that case. Note that the sample complexity in terms of number of trained tasks is not that high, even training on 20 randomly selected tasks leads to good generalization to the held-out tasks. Furthermore, the task embeddings generated by the system are organized by the features of the games, see appendix \ref{app_cards_tsne}. \par

@@

\paragraph{Meta-mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings (mappings over tasks) in order to flexibly reconfigure its behavior. For example, if the system is trained to map games to their losers variations, it can generalize this mapping to a game it has not been trained to map, even if the source or target of that mapping is held out from training. In Fig. \ref{cards_meta_map_results}) we demonstrate this by taking the mapped embedding and evaluating the reward received by playing the targeted game with it. This task is more difficult than simply learning to play a held out game from examples, because the system will actually receive no examples of the target game (when it is held out). Furthermore, in the case of the losers mapping, leaving the strategy unchanged would produce a large negative reward, and chance performance would produce 0 reward, so the results are quite good. \par

@@

%% TODO: language

\begin{figure}
\centering
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/language_meta_results.png}
\caption{From language input.}
\label{poly_meta_map_results_language}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/language_meta_mapping.png}
\caption{From language input.}
\label{cards_meta_map_results_language}
\end{subfigure}
\caption{The EML system can perform meta-mappings from language cues rather than meta-mapping examples, and can even generalize these. Compare to Fig. \ref{meta_map_results}, which shows the same results when using examples instead of language. (\subref{poly_meta_map_results_language}) In the polynomials domain, language cues still lead to good performance, even on held-out tasks or held-out meta-mappings, although examples perform slightly better (Fig. \ref{poly_meta_map_results_examples}). (\subref{cards_meta_map_results_language}) Similarly, in the cards domain, language cues perform well. Error-bars are bootstrap 95\%-CIs.}
\label{language_meta_map_results}
\end{figure}
\paragraph{Language:} Language is a fundamental piece of human flexibility, and we would be remiss to ignore it. In many of the intuitive illustrations we've given, the examples of the meta-mapping are implicit in your prior knowledge about the world. For example, ``try to lose at go'' does not give explicit examples of the ``lose'' meta-mapping, but rather relies on prior knowledge of what losing means. This is a much more efficient way to achieve zero-shot performance on a new losing task. In order to replicate this, we trained the EML system with both meta-mappings based on examples, and meta-mappings based on language. In the language-based meta-mappings, a language input identifying the meta-mapping (but not the basic task to apply it to) is provided to the network. This language input is encoded by a language encoder, and then provided as input to $\mathcal{H}$ instead of the output of $\mathcal{M}$. The meta-mapping then proceeds as normal --- $\mathcal{H}$ parameterizes $F$, which is used to transform the embedding of the input task to produce an embedding for the target task. \par
This language-cued meta-mapping approach also yields good performance (Fig. \ref{language_meta_map_results}). However, examples of the meta-mapping are slightly better, especially for meta-mappings not seen during training, presumably because examples provide a richer description. In Appendix \ref{app_why_meta_mapping} we show that using language to specify a meta-mapping performs better than using language to directly specify the target task, presumably by leveraging the richer task representation of the task embedding. \par

@@

To test this idea, we pre-trained the system on 100 polynomial tasks, and then introduced 100 new tasks. We trained on these new tasks by starting from the meta-network's ``guess'' at the correct task embedding, and then optimizing this embedding without altering the other parameters. The results are shown in Fig. \ref{poly_continual_results}. The meta-network embeddings offer good immediate performance, and substantially accelerate the optimization process, compared to a randomly-initialized embedding (see supp. Fig. \ref{supp_poly_continual_results} for a more direct comparison). Furthermore, this ability to learn is due to training, not simply the expressiveness of the architecture, as is shown by attempting the same with an untrained network. \par 

>>>>

Given this, the task of mapping inputs to outputs can be framed as trying to find a transformation of the representational space that takes the (embedded) inputs from the training set to embeddings that will decode to the target outputs. These transformations are performed by a system with the following components (see fig. \ref{architecture_inference_fig}): $\mathcal{M}: \{(Z, Z), ...\} \rightarrow Z $ -- the meta network, which collapses a dataset of (input embedding, target embedding) pairs to produce a single function embedding. $\mathcal{H}: Z \rightarrow \text{parameters}$ -- the hyper network, which maps a function embedding to parameters. $F: Z \rightarrow Z$ -- the transformation, implemented by a deep network parameterized by $\mathcal{H}$. \par

@@

%% Cut example if necessary
As a proof of concept, we first evaluated the system on the task of learning polynomials of degree $\leq 2$ in 4 variables (i.e. the task was to regress functions of the form $p: \mathbb{R}^4 \rightarrow \mathbb{R}$ where $p \in \mathcal{P}_2 \left(\mathbb{R}\right)$, though the model was given no prior inductive bias toward polynomial forms). For example, if $p(w,x,y,z) = x$, the model might see examples like $(-1,1,1,1; 1)$ and $(0.7, 2.1, 1.3, -4; 2.1)$, and be evaluated on its output for points like $(-1, -1.3, 0.5, 0.3)$. This yields an infinite family of base-level tasks (the vector space of all such polynomials), as well as many families of meta-mappings over tasks (for example, multiplying polynomials by a constant, squaring them, or permuting their input variables). This allows us to not only examine the ability of the system to learn to learn polynomials from data, but also to adapt its learned representations in accordance with these meta-tasks. Details of the architecture and training can be found in appendix \ref{app_detailed_methods}.\par
\vspace{-0.7em}
\paragraph{Basic meta-learning:} First, we show that the system is able to achieve the basic goal of learning a held-out polynomial from a few data points in fig. \ref{poly_basic_results} (with good sample-efficiency, see supp. fig. \ref{supp_poly_sweep_results}). \par 
\vspace{-0.7em}
\paragraph{Meta-mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings over polynomials in order to flexibly reconfigure its behavior, as shown in fig. \ref{poly_meta_map_results}. We train the system to perform a variety of mappings, for example switch the first two inputs of the polynomial, add 3 to the polynomial, or square the polynomial. We then test its ability to generalize to held-out mappings, for example a permutation of the polynomial inputs which it has not encountered before, or an additive shift in the polynomial which it has not been trained on. The system is both able to apply its learned meta-mappings to held-out tasks, and to apply meta-mappings it has not been trained on, simply by seeing examples of the mapping. \par

@@

\paragraph{Basic meta-learning:} First, we show that the system is able to achieve the basic goal of playing a held-out game from examples in fig. \ref{cards_basic_results}. We compare two different possible hold-out sets: 1) train on half the tasks at random, or 2) specifically hold out all the ``losers'' variations of the ``straight flush'' game. In either of these cases, the meta-learning system achieves well above chance performance (0) at the held out tasks, although it is slightly worse at generalizing to the targeted hold out, despite having more training tasks in that case. Note that the sample complexity in terms of number of trained tasks is not that high, even training on 20 randomly selected tasks leads to good generalization to the held-out tasks. Furthermore, the task embeddings generated by the system are organized by the features of the games, see appendix \ref{app_cards_tsne}. \par

@@

\paragraph{Meta-mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings (mappings over tasks) in order to flexibly reconfigure its behavior. For example, if the system is trained to map games to their losers variations, it can generalize this mapping to a game it has not been trained to map, even if the source or target of that mapping is held out from training. In fig. \ref{cards_meta_map_results}) we demonstrate this by taking the mapped embedding and evaluating the reward received by playing the targeted game with it. This task is more difficult than simply learning to play a held out game from examples, because the system will actually receive no examples of the target game (when it is held out). Furthermore, in the case of the losers mapping, leaving the strategy unchanged would produce a large negative reward, and chance performance would produce 0 reward, so the results are quite good. \par

@@

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/language_meta_results.png}
\caption{From language input.}
\label{poly_meta_map_results_language}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/language_meta_mapping.png}
\caption{From language input.}
\label{cards_meta_map_results_language}
\caption{The EML system can perform meta-mappings from language cues rather than meta-mapping examples, and can even generalize these. Compare to fig. \ref{meta_map_results}, which shows the same results when using examples instead of language. (\subref{poly_meta_map_results_language}) In the polynomials domain, language cues still lead to good performance, even on held-out tasks or held-out meta-mappings, although examples perform slightly better (fig. \ref{poly_meta_map_results_examples}). (\subref{cards_meta_map_results_language}) Similarly, in the cards domain, language cues perform well. Error-bars are bootstrap 95\%-CIs.}
\label{language_meta_map_results}
\paragraph{Language:} Language is a fundamental piece of human flexibility, and we would be remiss to ignore it. In many of the intuitive illustrations we've given, the examples of the meta-mapping are implicit in your prior knowledge about the world. For example, ``try to lose at go'' does not give explicit examples of the ``lose'' meta-mapping, but rather relies on prior knowledge of what losing means. This is a much more efficient way to achieve zero-shot performance on a new losing task. In order to replicate this, we trained the EML system with both meta-mappings based on examples, and meta-mappings based on language. In the language-based meta-mappings, a language input identifying the meta-mapping (but not the basic task to apply it to) is provided to the network. This language input is encoded by a language encoder, and then provided as input to $\mathcal{H}$ instead of the output of $\mathcal{M}$. The meta-mapping then proceeds as normal --- $\mathcal{H}$ parameterizes $F$, which is used to transform the embedding of the input task to produce an embedding for the target task. \par
This language-cued meta-mapping approach also yields good performance (fig. \ref{language_meta_map_results}). However, examples of the meta-mapping are slightly better, especially for meta-mappings not seen during training, presumably because examples provide a richer description. In Appendix \ref{app_why_meta_mapping} we show that using language to specify a meta-mapping performs better than using language to directly specify the target task, presumably by leveraging the richer task representation of the task embedding. \par

@@

To test this idea, we pre-trained the system on 100 polynomial tasks, and then introduced 100 new tasks. We trained on these new tasks by starting from the meta-network's ``guess'' at the correct task embedding, and then optimizing this embedding without altering the other parameters. The results are shown in fig. \ref{poly_continual_results}. The meta-network embeddings offer good immediate performance, and substantially accelerate the optimization process, compared to a randomly-initialized embedding (see supp. fig. \ref{supp_poly_continual_results} for a more direct comparison). Furthermore, this ability to learn is due to training, not simply the expressiveness of the architecture, as is shown by attempting the same with an untrained network. \par 
